Notes about Periodic table for topological insulators and superconductors

### 2016-11-02 Notes of various papers/Kitaev. Periodic table for topological insulators and superconductors/notes.tex
New notes to the group theory - representation.

\usepackage{float}  % For H in figures
\usepackage{amsmath} % For math
\usepackage{amssymb}
\usepackage{bbm} % for numbers within mathbb
\usepackage{mathrsfs} % For \mathscr{ABC}
% Followings are for the special character: differential "d".
\newcommand*\diff{\mathop{}\!\mathrm{d}}
\newcommand*\Diff[1]{\mathop{}\!\mathrm{d^#1}}
\numberwithin{equation}{subsection} % have the enumeration go to the subsection level.
                                    % See:https://en.wikibooks.org/wiki/LaTeX/Advanced_Mathematics
\usepackage{graphicx}   % need for figures
\usepackage{cite} % need for bibligraphy.
\usepackage[unicode]{hyperref}  % make every cite a link
\usepackage{CJKutf8} % For Chinese characters
\usepackage{fancyref} % For easy adding figure,equation etc in reference. Use \fref or \Fref instead of \ref

% Following is for theorems etc environments
% http://tex.stackexchange.com/questions/45817/theorem-definition-lemma-problem-numbering && https://en.wikibooks.org/wiki/LaTeX/Theorems
\usepackage{amsthm}
\newtheorem{defi}{Definition}[section]
\newtheorem{thm}{Theorem}[section]
\newtheorem{lemma}{Lemma}[section]
\newtheorem{remark}{Remark}[section]
\newtheorem{prop}{Proposition}[section]
\newtheorem{coro}{Corollary}[section]
\newtheorem{fact}{Fact}[section]
\theoremstyle{definition}
\newtheorem{ex}{Example}[section]
\newtheorem{argument}{Argument}[section]

% A list of nomenclatures.
\usepackage{nomencl}
\makenomenclature

% For drawing diagrams with arrows
\usepackage[all]{xy}

% My own physics package
% The following line load the package xparse with additional option to
% prevent the annoying warnings, which are caused by the package
% "physics" loaded in package "physics-taper".
\usepackage[log-declarations=false]{xparse}
\usepackage{physics-taper}

\title{Notes about Spin and TR Symmetry}
\date{\today}
\author{Taper}

@@

\maketitle
\abstract{
This is a note about TR symmetry, with its application in the theory of
condensed matter physics. The two reference book are the classic by
Sakurai \cite{sakurai}, and the other by Bernevig \cite{bernevig}.
It contains a summary of treatment of Spin (actually, rotation) in
quantum mechanics, which is needed to discuss the TR symmetry operator.
}
\tableofcontents

\section{Review of Spin theory}
\label{sec:Review-of-Spin-theory}
Outline:
\begin{enumerate}
    \item The representation of infinitesimal operator.
    \item From traditional rotational operators to commutation
        relationship
    \item From commutators to their eigenvalues.
    \item Summarize some further results regarding the matrix
        representation of those operators.
\end{enumerate}

\subsection{The representation of transformation}
\label{sec:The-representation-of-transformation}

To find an operator for a physical transformation, we naturally encounters
the theory of representation of groups. Why groups? See the following
notes for an bird view. (The note is an excerpt from my notes about
Quantum Field Theory by Weinberg, which is still in its infancy because
the book is so hard to read)
\begin{quote}
    The set of all symmetries transformations obviously can have a group
    structure. By giving each symmetry transformation a unitary or
    antiunitary operator, a representation of such group is obtained.
    However, such a representation can be projective (projective as in
    projective geometry, or $\mathbb{CP}^n$), since operators acts on ket
    spaces, which is already projective (i.e. within a freedom of phase). 
\end{quote}

Here I outlined one general procedure to obtain such representation from
infinitesimal transformations. This procedure is discussed in chapter 1
and 2 of \cite{sakurai}.
% TODO update more specific location inside that book.
We first needs two thing to characterize a tranformation: a Hermitian
operator $G$ and an infinitesimal change $\varepsilon$. I am not sure about
how to choose $G$, but I think the choice for $\varepsilon$ should be
% TODO how to choose G, and what is G
clear. Here are three typical examples.

For displacement in $x$ direction:
\begin{align}
    G \rightarrow \frac{p_x}{\hbar},\, \varepsilon \rightarrow \diff x
\end{align}
For an infinitesimal time evolution:
\begin{align}
    G \rightarrow \frac{H}{\hbar},\, \varepsilon \rightarrow \diff t
\end{align}
For infinitesimal rotation around $k$ th axis:
\begin{align}
    G \rightarrow \frac{J_k}{\hbar},\, \varepsilon \rightarrow \diff \phi
    \label{eq:revSpin.repreTrans.angulalMomGen}
\end{align}
Or more general (let the rotation be characterized by $\mathbf{J}\cdot
\hat{\mathbf{n}}$, where $\hat{\mathbf{n}}$ is a unit vector), we have:
\begin{align}
    G \rightarrow \frac{\mathbf{J}\cdot \hat{\mathbf{n}} }{\hbar},\,
    \varepsilon \rightarrow \diff \phi
\end{align}

Here $\mathbf{J}=(J_x,J_y,J_z)$ is defined such that the substitution
\ref{eq:revSpin.repreTrans.angulalMomGen} indeed produces a rotation.
What do we mean by "indeed" here? This will be explained in section
\ref{sec:From-classical-mechanic-to-commutators}.
\label{page:question-def_of_J}

Then the operator corresponding to the infinitesimal transformation is
\begin{align}
    U_\varepsilon= 1-i\, G\varepsilon
\end{align}
And for a finite transformation (let $\Delta$ denotes the finite
transformation):
\begin{align}
    \lim_{N\to \infty}\left[ 1-iG\frac{\Delta}{N}\right]^N =
    \exp(-iG\Delta)
\end{align}
For example, in the case of rotation we have (using $\mathscr{D}$ to
denote the operator):
\begin{align}
    \mathscr{D}(\hat{\mathbf{n}},\diff \phi) =
    1-i\frac{\mathbf{J}\cdot\hat{\mathbf{n}}}{\hbar}\diff\phi
    \\
    \mathscr{D}(z,\phi)=\exp(\frac{-iJ_z \phi}{\hbar})
    \label{eq:revSpin.repreTrans.repreFiniteRot}
\end{align}

Note that, here $G$ is regarded as a generator of physical transformation:
\begin{table}[H]
    \centering
    \caption{Generators}
    \begin{tabular}{|c|c|}
    \hline
    Generator                     & Physically generated transformation \\
    \hline
    Momentum $p_x$                & Translation $\diff x$ \\
    \hline
    Energy $H$                    & Time evolution: $\diff t$ \\
    \hline
    Angular momentum $J_x$ & Rotation about $x$ axis \\
    \hline
    \end{tabular}
\end{table}
\subsection{From classical mechanics to commutators in quantum mechanics}
\label{sec:From-classical-mechanic-to-commutators}

This part is discussed in secion 3.1 of \cite{sakurai}.

The commutation relationship is obtained by considering infinitesimal
transformation. In general, classical rotation can be decomposed into
rotation about three axes. We can use linear transformation to express
them. For example, rotation about $z$ axis is:
\begin{align}
    R_z(\phi) = \left( \begin{array}{ccc}
         \cos (\phi ) & -\sin (\phi ) & 0 \\
         \sin (\phi ) & \cos (\phi ) & 0 \\
         0 & 0 & 1 \\ \end{array} \right)
\end{align}
Put it into infinitesimal amount, one has:
\begin{align} 
    R_z(\varepsilon)=\left( \begin{array}{ccc}
     1-\frac{\varepsilon ^2}{2} & -\varepsilon               & 0 \\
     \varepsilon                & 1-\frac{\varepsilon ^2}{2} & 0 \\
     0                          & 0                          & 1 \\
\end{array} \right) \end{align}
Similar procedure could be done for other axes (see equation 3.1.4 and
3.1.5 in \cite{sakurai}). By calculation one finds that
$$ R_x(\varepsilon)R_y(\varepsilon)=R_y(\varepsilon)R_x(\varepsilon)$$
if terms higher than $\varepsilon$ is ignored. This means that
infinitesimal rotation $\diff \mathbf{\omega}$ can be safely considered as
vectors.

However if only terms of order equal and lower than $\varepsilon^2$ is kept
(note especially that $\varepsilon^4$ is ignored), one finds:
\begin{align}
    \label{eq:revSpin.fromCM2QM.classicalRotComu}
    R_x(\varepsilon)R_y(\varepsilon)-R_y(\varepsilon)R_x(\varepsilon)=
    R_z(\varepsilon^2)-\mathbb{I}
\end{align}

Expand equation \ref{eq:revSpin.repreTrans.repreFiniteRot} in orders of
$\phi$ and keep only terms with order $leq \phi^2$, we have:
$$
\mathscr{D}(x,\varepsilon)=
1-\frac{iJ_x \varepsilon}{\hbar}-\frac{J_x^2 \varepsilon^2}{2\hbar^2}
$$
and so on for other axes.

Plug these results into the classical commutation relationship
\ref{eq:revSpin.fromCM2QM.classicalRotComu}, and with some calculation,
one will find:
\begin{align}
    \label{eq:revSpin.fromCM2QM.comuOfAngularM}
    [J_i,J_j]=i\hbar \varepsilon_{ijk}J_k
\end{align}
here $\varepsilon_{ijk}$ is the Levi-Civita notation in $3$ dimension.
This relation also ensures that the operator $\mathbf{J}$ we defined
indeed corresponds to a physical rotation, which answers the question is
mentioned in page \pageref{page:question-def_of_J}.

\subsection{From commutators to eigenvalues}
\label{sec:From-commutators-to-eigenvalues}
Now one may use the commutation relation
\ref{eq:revSpin.fromCM2QM.comuOfAngularM} to get the eigenvalues of these
operators. The procedure is quite standard and can be found in section 3.5
of \cite{sakurai}. There the author defines the ladder operators:
\begin{defi}[Ladder operator]
\nomenclature{Ladder operator}{\nomrefpage.}
    \begin{align}
        J_{\pm}\equiv J_x \pm i J_y
    \end{align}
\end{defi}
Then he derives the cummutation relationship:
\begin{align}
    [\mathbf{J}^2,J_k] &= 0\\
    [J_+,J_-] &=2\hbar J_z \\
    [J_z,J_\pm] &= \pm \hbar J_\pm \\
    [\mathbf{J}^2,J_\pm] &=0 \\
\end{align}
Chossing the $z$ direction as eigenstate. Denote the simultaneous
eigenstates of $\mathbf{J}^2$ and $J_z$ as $\ket{j,m}$, such that:
\begin{align}
    \mathbf{J}^2\ket{j,m} &= j(j+1)\hbar^2 \ket{j,m} \\
    J_z \ket{j,m} &= m \hbar\ket{j,m}
\end{align}
(such a strange choice is conventional in the case of spin).

As usual, one finds that
\begin{itemize}
    \item For $J_z$, $J_\pm$ will pull up/lower its eigenvalue for
        eigenstate in units of $\hbar$.
    \item For $\mathbf{J}^2$, $J_\pm$ will not alter its eigenvectors,
        since it commutes with $\mathbf{J}^2$.
\end{itemize} 
By above relationships, one can derive the following facts (for
convenience let $a$ be the eigenvalue for $\mathbf{J}^2$, and $b$ be the
eigenvalue for $J_z$.):
\begin{itemize}
    \item $a\geq b^2$, i.e $b$ is bounded by $a$. More specifically,
    \item $a=b_{\text{max}}(b_{\text{max}}+\hbar)$.
    \item Also, $-b_{\text{max}}\leq b \leq b_{\text{max}}$.
    \item Using $b_{\text{max}}=b_{\text{min}}+n\hbar$ ($n$ is some
        integer), one gets
    \item $b_{\text{max}}= \frac{n\hbar}{2}$. This, by identifying
        $j=\frac{n}{2}$, one has
\end{itemize}
\begin{align}
    a &= j(j+1)\hbar^2 \\
    b &= m\hbar\\
    (m &= -j, -j+1,\cdots, j-1,j;\, j=\frac{n}{2}) \nonumber
\end{align}
\subsection{Getting the matrix of rotation operators}
\label{sec:Using-Euler-angles-to-get-matrix-of-rotation-operators}

Simply by commutation relationships above, one can easy get:
\begin{align}
    \bra{j',m'}\mathbf{J}^2\ket{j,m} &= 
        j(j+1)\hbar^2 \delta_{j'j}\delta_{m'm} \\
    \bra{j',m'}J_z\ket{j,m} &= 
        m\hbar \delta_{j'j}\delta_{m'm} \\
    \bra{j',m'}J_\pm\ket{j,m} &= 
        \sqrt{(j\mp m)(j\pm m+1)} \hbar \delta_{j'j}\delta_{m',m\pm 1}
        \label{eq:revSpin.getMatRot.J_pm}
\end{align}

To get matrix element for
$\exp(\frac{-i\mathbf{J}\cdot\hat{\mathbf{n}}\phi}{\hbar})$, one needs to
use Euler angles.

    \subsubsection{Eulerian Angles}
    \label{sec:Eulerian-Angles}
    The Euler angles is discussed in section 3.3 of \cite{sakurai}.
    However, Sakurai's usage in his book \cite{sakurai} is a bit different
    from the usual one. But the following graph taken from that section
    should explains all:
    \begin{figure}[H]
        \centering
        \includegraphics[width=1.2\linewidth]{pics/{fig3.4in-Sakurai-Euler-angles}.png}
        \caption{Euler Angles (fig 3.4 in that book)}
        \label{fig:fig3.4-Sakurai-Euler-angles}
    \end{figure}

    One can find that a Euler rotaion characterized by
    $(\alpha,\beta,\gamma)$ could be decomposed into:
    \begin{align}
        R(\alpha,\beta,\gamma)=R_z(\alpha)R_y(\beta)R_z(\gamma)
    \end{align}

    \subsubsection{Using Euler angles to get matrix for rotation operator}
    \label{sec:Using Euler angles to get matrix for rotation operator}

    As the title suggests, in section 3.5 of \cite{sakurai}, we have
    $$
    \exp\left(\frac{-i\mathbf{J}\cdot\hat{\mathbf{n}}\phi}{\hbar}\right)
    =
    \exp\left(\frac{-i J_z\alpha}{\hbar}\right)
    \exp\left(\frac{-i J_y\beta}{\hbar}\right)
    \exp\left(\frac{-i J_z\gamma}{\hbar}\right)
    $$,
    define 
    \begin{align}
        \nomenclature{
            Wigner function $\mathscr{D}^{(j)_{m'm}}$}{\nomrefpage.}
        \mathscr{D}^{(j)}_{m'm}\equiv
        \bra{j,m'}
            \exp\left(\frac{-i J_z\alpha}{\hbar}\right)
            \exp\left(\frac{-i J_y\beta}{\hbar}\right)
            \exp\left(\frac{-i J_z\gamma}{\hbar}\right)
        \ket{j,m}
    \end{align}
    This $\mathscr{D}$ is sometimes called the \textbf{Wigner function}.
    Note that we have implicitly assumed that the rotation in 
    $\exp\left(\frac{-i\mathbf{J}\cdot\hat{\mathbf{n}}\phi}{\hbar}\right)$
    does not alter the eigenstate of length $\mathbf{J}^2$. This is
    physically satisfying, and can also be proved (equation 3.5.43 in
    \cite{sakurai}).

    % TODO I thin we should distinguish of two concepts of rotation:
    % J and exp(-iJ/h).

    Since we have diagonalize using the eigenstate of $J_z$, we have:
    \begin{align}
        \mathscr{D}^{(j)}_{m'm} =
            e^{-i(m'\alpha+m\gamma)}\bra{j,m'}
                \exp\left(\frac{-i J_y\beta}{\hbar}\right)
            \ket{j,m}
    \end{align}
    The remaining part can only be calculated case by case. Denote it as 
    $d^{(j)}_{m'm}$\nomenclature{ $d^{(j)}_{m'm}$}{\nomrefpage.}, i.e.:
    \begin{align}
        d^{(j)}_{m'm}\equiv \bra{j,m'}
                \exp\left(\frac{-i J_y\beta}{\hbar}\right)
            \ket{j,m}
    \end{align}

    Detailed calculation could start from equation
    \ref{eq:revSpin.getMatRot.J_pm}.
    In case of $j=\frac{1}{2}$, as is familiar for use, it is:
    \begin{align}
        d^{\frac{1}{2}}=\left( \begin{array}{cc}
            \cos \left(\frac{\beta }{2}\right) 
                & -\sin \left(\frac{\beta}{2}\right) \\
            \sin \left(\frac{\beta }{2}\right) 
                & \cos \left(\frac{\beta}{2}\right) \\
            \end{array} \right)
    \end{align}
    Footnote \footnote{This can also be obtained using the Pauli matrices,
    which can be found in my miscellaneous notes.}

    \subsubsection{Pauli matrices in half-spin: about \texorpdfstring{$e^{-i\pi S_y/\hbar}$}{}}
    \label{sec:About-exp(-i*pi-S_y/hbar)}

    Since everyone is familiar with the Pauli matrices in spin $1/2$
    system, I here only shows a useful result. This result is related to
    equation (4.4.65) in \cite{sakurai}.

    Since 
    $$S_y=\frac{\hbar}{2}\sigma_y=\frac{\hbar}{2}
        \left( \begin{array}{cc}
         0 & -i \\
         i & 0 \\
        \end{array} \right)$$
    It is easy to check that $\sigma_y^2 = \mathbb{I}$. So:
    \begin{align*}
        e^{-i\phi S_y/\hbar}=
        e^{-i\frac{\phi}{2}\sigma_y}&=\sum_{n=0}^{\infty}
         \frac{(-i\frac{\phi}{2})^n(\sigma_y)^n}{n!}\nonumber\\
         &=\sum_{\text{odd }n}-\frac{i(\frac{\phi}{2})^n}{n!}\sigma_y
           + \sum_{\text{even }n}\frac{(i\frac{\phi}{2})^n}{n!}\mathbb{I}
    \end{align*}
    Using the taylor expansions:
    \begin{align}
        \sinh{x}&= \sum_{n=0}^{\infty}\frac{x^{2n+1}}{(2n+1)!}
        \text{ (odd terms)}\\
        \cosh{x}&= \sum_{n=0}^{\infty}\frac{x^{2n}}{(2n)!}
        \text{ (even terms)}
    \end{align}
    We see that:
    \begin{align}
        e^{-i\phi S_y/\hbar}=e^{-i\frac{\phi}{2}\sigma_y}&=
         - \sigma_y\sinh{i\frac{\phi}{2}}+ \mathbb{I}\cosh{i\frac{\phi}{2}}
         \nonumber\\
         &=
         -i\sigma_y \sin{\frac{\phi}{2}}+\mathbb{I}\cos{\frac{\phi}{2}}
        \\
         &=-i\frac{2}{\hbar}S_y\sin{\frac{\phi}{2}}
            +\mathbb{I}\cos{\frac{\phi}{2}}
        \\
         &=\left( \begin{array}{cc}
            \cos(\frac{\phi }{2}) & -\sin(\frac{\phi }{2}) \\
            \sin (\frac{\phi }{2}) & \cos (\frac{\phi }{2}) \\
         \end{array}\right)
    \end{align}
    So
    $$e^{-i\pi S_y/\hbar} = -i\frac{2S_y}{\hbar}$$
    By exactly the same argument, or simply by the symmetry
    ($x\leftrightarrow y$), we
    have:
    \begin{align}
        e^{-i\phi S_x/\hbar}=e^{-i\frac{\phi}{2}\sigma_x}&=
         -i\sigma_x \sin{\frac{\phi}{2}}+\mathbb{I}\cos{\frac{\phi}{2}}
        \\
        &=-i\frac{2}{\hbar}S_x\sin{\frac{\phi}{2}}
            +\mathbb{I}\cos{\frac{\phi}{2}}
        \\
        &=\left(\begin{array}{cc}
            \cos(\frac{\phi }{2}) & -\sin(\frac{\phi }{2}) \\
            \sin (\frac{\phi }{2}) & \cos(\frac{\phi }{2}) \\
         \end{array}\right)
    \end{align}
    Also, one can easily find that:
    \begin{align}
        e^{-i\phi S_z/\hbar}=e^{-i\frac{\phi}{2}\sigma_z}&=
          \left(\begin{array}{cc}
              e^{-i\frac{\phi }{2}} & 0 \\
              0                     & e^{i\frac{\phi }{2}} \\
         \end{array}\right)
    \end{align}
    Also, we have a formula that encompass all the above relationship
    (proved in page 170, section 3.2 of \cite{sakurai}):
    \begin{align}
        \exp(\frac{-i\vec{\sigma}\cdot\hat{\vec{n}}\phi}{2})
        =
        \cos(\frac{\phi}{2})\mathbb{I}-i\vec\sigma\cdot\hat{\vec{n}}
        \sin(\frac{\phi}{2})
    \end{align}
    where $\vec\sigma=(\sigma_x,\sigma_y,\sigma_z)$, and 
    $\hat{\vec{n}}$ is any unit vector.

    \subsubsection{Eigenstate for \texorpdfstring{$\mathbf{S}\cdot
    \hat{\mathbf{n}}$}{}}
    \label{sec:Eigenstate_for_Sn}
    Here is a physical arguemnt to construct an eigenstate for the
    operator $\mathbf{S}\cdot \hat{\mathbf{n}}$. It mimics what was done
    in page 172 of \cite{sakurai}. This is certainly correct when
    $j=\frac{1}{2}$. However, I am not sure about its validity in case of
    higher spin.
    % TODO check this!

    Suppose we have a state of the highest eigenvalue for $J_z$, i.e.
    a state labeled by $\ket{j,j}$, then in principle any state can
    be obtained from this state after a rotatation and a scale. But since
    $\ket{\mathbf{n}}$ is unit vector, it can be characterized by two
    angle: the polar angle $\beta$ and the azimuthal angles $\alpha$.
    As shown in the following picture:
    \begin{figure}[H]
        \centering
        \includegraphics[width=0.8\linewidth]{pics/{fig3.3in-Sakurai-polar-n-azimuthal}.png}
        \caption{Polar and azimuthal angle (fig 3.3 in \cite{sakurai})}
    \end{figure}

    So I think the eigenstate should be:
    \begin{align}
        \label{eq:revSpin.getMatRot.Sn.Snket}
        \ket{\hat{\mathbf{n}};+}= e^{-iS_z \alpha/\hbar}
        e^{-iS_y \beta/\hbar} \ket{+}
    \end{align}
    here $\ket{+}$ denotes the state $\ket{j,j}$. The eigenvalue is of
    course $j\hbar$.

    This is confirmed when $j=\frac{1}{2}$ in page 171 of \cite{sakurai},
    but not in other cases.

\section{Review of Symmetry Operator}
\label{sec:Review-of-Symmetry-Operator}

This part is discussed in section 4.4 of \cite{sakurai}.

Outline:
\begin{enumerate}
    \item A note about antilinear operator, and we should treat it with
        care.
    \item The invariance of probability and Wigner's theorem
    \item The values of measurement before and after transformation
\end{enumerate}

\subsection{A note about antiunitary operator}
\label{sec:A-note-about-antiunitary-operator}
Please note that the Dirac bra-ket notation is only convenient when we are
dealing with unitary operators. When antiunitary operators, or arbitrary
operators are considered, the situation can get tricky. Here are two
examples.

    \subsubsection{Formula for base transformation}
    \label{sec:Formula-for-base-transformation}

    Suppose we transform the basis
    $\ket{e_i}$ by an operator $\mathcal{O}$, what is the formula for the
    corresponding basis for the dual space? The requirement of dual basis
    requires that:
    $$ \bra{\widetilde{e}_i} = \bra{\widetilde{e}_i}U^{-1}$$
    But by the "dual-correspondance" in Dirac notation, we expect:
    $$ \bra{\widetilde{e}_i} = \bra{\widetilde{e}_i}U^{\dagger}$$
    The two notion coincide only when $U$ is unitary.

    Similarly, consider the operator $\ket{e_i}\bra{e_i}$, if we perform a
    scale of the basis $\ket{e_i}\to \lambda\ket{e_i}$, then how will the
    above operator transform? Is it
    $$ \ket{\lambda e_i}\bra{\lambda e_i}$$
    or is it:
    $$ \ket{\lambda^{-1} e_i}\bra{\lambda e_i}$$
    Of course, when the transformation is unitary, the two notion
    coincide, because $\lambda^*=\lambda^{-1}$ implies
    $\lambda=\lambda^{-1}$.
    \subsubsection{The adjoint of antilinear operator}
    label{sec:The-adjoint-of-antilinear-operator}

    If we were to define adjoint of an antilinear operator, we might
    run into troubles with traditional definition. For example, suppose
    $A$ is antilinear, let us multiply it with $c\mathbbm{1}$, where $c$
    is just some nonzero constant.

    We have:
    \begin{align}
        \braket{\phi|cA \psi} = c\braket{\phi|A\psi}
    \end{align}
    on the other hand:
    \begin{align}
        &\braket{\phi|cA \psi} = \braket{cA\psi|\phi}^*
            = \braket{\psi|A^\dagger c^* \phi}^* 
            =(c\braket{\psi|A^\dagger \phi})^* \nonumber\\
            =& c^* \braket{A\psi|\phi}^* 
            = c^* \braket{\phi|A\psi}
    \end{align}
    A contracdiction. Then, it was proposed, in the first answer to
    this Math.SE post\cite{math.SE-adjoint-of-antilinear}, that we define
    the adjoint operator differently for antilinear operator $A$:
    \begin{align}
        \braket{A^\dagger v|w} = \braket{v| Aw}^*
    \end{align}
    Then the above discrepancy will not exist anymore (easily confirmed).

    Here I also include the complete answer for reference:
    \begin{center}\noindent\rule{8cm}{0.4pt}\end{center}
    \begin{quote}
        I) First of all, one should never use the [Dirac bra-ket notation](http://en.wikipedia.org/wiki/Bra-ket\_notation) (in its ultimate version where an operator acts to the right on kets and to the left on bras) to consider the definition of [adjointness](http://en.wikipedia.org/wiki/Adjoint\_operator), since the notation was designed to make the adjointness property look like a mathematical triviality, which it is not. See also [this](http://physics.stackexchange.com/q/43069/2451) Phys.SE post.

        II) OP's question(v1) about the existence of the adjoint of an [antilinear](http://en.wikipedia.org/wiki/Antilinear\_map) operator is an interesting mathematical question, which is rarely treated in textbooks, because they usually start by assuming that operators are $\mathbb{C}$-linear.

        III) Let us next recall the mathematical definition of the adjoint of a linear operator. Let there be a [Hilbert space](http://en.wikipedia.org/wiki/Hilbert\_space) $H$ over a [field](http://en.wikipedia.org/wiki/Field\_\%28mathematics\%29) $\mathbb{F}$, which in principle could be either real or complex numbers, $\mathbb{F}=\mathbb{R}$ or $\mathbb{F}=\mathbb{C}$. Of course in quantum mechanics, $\mathbb{F}=\mathbb{C}$. In the complex case, we will use the standard physicist's convention that the [inner product/sequilinear form](http://en.wikipedia.org/wiki/Sesquilinear\_form) $\langle \cdot | \cdot \rangle$ is conjugated $\mathbb{C}$-linear in the first entry, and $\mathbb{C}$-linear in the second entry.

        Recall [Riesz' representation theorem](http://en.wikipedia.org/wiki/Riesz\_representation\_theorem): For each continuous $\mathbb{F}$-linear functional $f: H \to \mathbb{F}$ there exists a unique vector $u\in H$ such that
        $$\text{tag\{1\}} f(\cdot)~=~\langle u | \cdot \rangle.$$

        Let $A:H\to H$ be a continuous$^1$ $\mathbb{F}$-linear operator. Let $v\in H$ be a vector. Consider the continuous $\mathbb{F}$-linear functional

        $$\text{tag\{2\}} f(\cdot)~=~\langle v | A(\cdot) \rangle.$$

        The value $A^{\dagger}v\in H$ of the adjoint operator $A^{\dagger}$ at the vector $v\in H$ is by definition the unique vector $u\in H$, guaranteed by Riesz' representation theorem, such that 
        $$\text{tag\{3\}} f(\cdot)~=~\langle u | \cdot \rangle.$$

        In other words, 
        $$\text{tag\{4\}} \langle A^{\dagger}v | w \rangle~=~\langle u | w \rangle~=~f(w)=\langle v | Aw \rangle. $$

        It is straightforward to check that the adjoint operator $A^{\dagger}:H\to H$ defined this way becomes an $\mathbb{F}$-linear operator as well.

        IV) Finally, let us return to OP's question and consider the definition of the adjoint of an antilinear operator. The definition will rely on the complex version of Riesz' representation theorem. Let $H$ be given a complex Hilbert space, and let  $A:H\to H$ be an antilinear continuous operator. In this case, the above equations (2) and (4) should be replaced with

        $$\text{tag\{2'\}} f(\cdot)~=~\overline{\langle v | A(\cdot) \rangle},$$

        and

        $$\text{tag\{4'\}} \langle A^{\dagger}v | w \rangle~=~\langle u | w \rangle~=~f(w)=\overline{\langle v | Aw \rangle}, $$

        respectively. Note that $f$ is a $\mathbb{C}$-linear functional.

        It is straightforward to check that the adjoint operator $A^{\dagger}:H\to H$ defined this way becomes an antilinear operator as well.

        --

        $^{1}$We will ignore subtleties with [discontinuous/unbounded operators](http://en.wikipedia.org/wiki/Unbounded\_operator), domains, [selfadjoint extensions](http://en.wikipedia.org/wiki/Extensions\_of\_symmetric\_operators), etc., in this answer.
    \end{quote}

    \begin{center}\noindent\rule{8cm}{0.4pt}\end{center}

    \subsubsection{Conclusion}
    \label{sec:Conclusion}
    For antilinear operators, we should always working with it acting on
    the kets. See page 289 in section 4.4 of \cite{sakurai}.

\subsection{Properties of a symmetry operator}
\label{sec:Property-of-symmetry-operator}
Consider a symmetry operation acting on states such that:
\begin{align}
    \ket{\alpha}\to \ket{\widetilde{\alpha}}\,\,
    \ket{\beta}\to \ket{\widetilde{\beta}}
\end{align}
We require that the probability, the most important thing in quantum
mechanics, is unchaged:
\begin{align}
    |\braket{\widetilde{\beta}|\widetilde{\alpha}}| =
        |\braket{\beta|\alpha}|
\end{align}
Wigner has proved that\footnote{
Though it was only mentioned in page 289 of \cite{sakurai}.},
all such symmetry transformations can be characterized either by an unitary
or an antiunitary operator. Also, any antiunitary operator $\theta$ can be
expressed as
\begin{align}
    \theta = U K
\end{align}
where $U$ is unitary, and $K$ is complex conjugation. Three facts about
$K$:
\begin{itemize}
    \item \textit{"acts only on any coefficient that multiplies a ket
        (and stands on the right of K)."} 
    \item $K$ also does not act on the base ket, i.e. $K\ket{e}=\ket{e}$.
    \item $K$'s action depends heavily the base ket. This is obvious from
        the above two points. Of course, this is not important when every
        physical state is only projective. However, I still doubt whether
        this will call additional problem.
        \footnote{Consider $K\ket{e}=\ket{e}$, but when
        $\ket{e}=i\ket{e'}$, then $K\ket{e}=-i\ket{e'}=-\ket{e}$.}
\end{itemize}

\subsection{States and Operators after transformation}
\label{sec:States-and-Operators-after-transformation}
Let $U$ be an unitary operator and $A$ be an antiunitary operator. Suppose
we have $\ket{\alpha}$ and $\ket{\beta}$, and regard both to be
transformed under $U$ and $A$. Let:
\begin{align}
    \ket{\alpha_u}=U\ket{\alpha},\, \ket{\alpha_a}=A\ket{\alpha} \\
    \ket{\beta_u}=U\ket{\beta},\, \ket{\beta_a}=A\ket{\beta}
\end{align}
Then by defintion of unitary and antiunitary, we have:
\begin{align}
    \braket{\beta_u|\alpha_u} &=\braket{\beta|\alpha} \\
    \braket{\beta_a|\alpha_a}^* &=\braket{\beta|\alpha}
\end{align}

For any operator $\mathcal{O}$, we have, after transformation:
\begin{align}
    \mathcal{O}_u &= U\mathcal{O}U^\dagger \\
    \mathcal{O}_a &= A \mathcal{O}^\dagger A^{-1}
\end{align}
The two equation are mean to expresses the following idea:
\begin{align}
    \bra{\alpha}\mathcal{O}\ket{\beta} &=
    \braket{\widetilde{\alpha}|\mathcal{O}_u |\widetilde{\beta}} \\
    \bra{\alpha}\mathcal{O}\ket{\beta} &=
    \braket{\widetilde{\beta}|\mathcal{O}_a |\widetilde{\alpha}}
\end{align}

Proof may be found on page 292 of \cite{sakurai}.
\section{TR operator}

Finally I can turn to the TR operator. But before the formal discussion of
time reversal, I should first note that the name \textbf{time reversal} is
actually a misnomer. As mentioned in page 284, section 4.4 in 
\cite{sakurai}:
\begin{quote}
    What we do in this section can be more appropriately characterized by
    the term \textbf{reversal of motion}. Indeed, that is the phrase used
    by E.  Wigner, who formulated time reversal in a very fundamental
    paper written in 1932.
\end{quote}
This section can be found in section 4.4 of \cite{sakurai}.

>>>>

% My own physics package
% The following line load the package xparse with additional option to
% prevent the annoying warnings, which are caused by the package
% "physics" loaded in package "physicist-taper".
\usepackage[log-declarations=false]{xparse}
\usepackage{physicist-taper}

\makenomenclature

\title{Notes about Periodic table for topological insulators and superconductors}
\date{\today}
\author{Taper}

@@

\maketitle
\abstract{
As title suggests.
}
\tableofcontents
From the start: for \textbf{Gapped free-fermion} system, he mentions
the classification hitherto available. See
Table~\ref{tab:classification}.
\begin{table}[ht]
    \centering
    \caption{Classification}
    \label{tab:classification}
    \vspace{+0.5pt}
    \rowcolors{1}{white}{lightgrey}
    \tabulinesep=1.2mm
    \begin{tabu}{  X[l] X[l] X[l] X[l] }
        \textit{Symmetry} & \textit{Spatial Dimension} &
        \textit{Result} & \textit{Other Keywords} \\
        \hline
        T & 0 & An intger: the number of particle-occupied Kramers
        doublet states & \\
        T & 1 & None & \\
        T & 2 & $\mathbb{Z}_2$ & \\
        T & 3 & $\mathbb{Z}_2$ & $3D$ crystals have additional
        $3\mathbb{Z}_2$ invariant$ \Rightarrow$ "weak topological
        insulators \\
        Q(?) & 2 & Characterized by $\mu$ in units of $e^2/h$ & TKNN \\
        Q(?) & even  d & Topological invariant ($k$-th Chern number) \\
        Q(?) & 0 &  number of single-particle states with negative
energy ($E< E_F = 0$), which are filled with electrons. \\
        T\& Q \\
        No T \& No Q & 0 & $\mathbb{Z}_2$ \\
        No T \& No Q & 1 &$\mathbb{Z}_2$ &"majorana chain"  \\
        No T \& No Q & 2 & Topological number is integer. & Even-odd
        effects. \\
    \bottomrule
    \end{tabu}
\end{table}

\paragraph{Robustness in classification}
\begin{quote}
    ..., the topological numbers are insensitive to disorder and can even
    be defined without the spectral gap assumption, provided the
    eigenstates are localized.
\end{quote}

\paragraph{Equivalence of Hamiltonian} 
\begin{quote}
    ... Two Hamiltonians belong to the same phase if they can be
    continuously transformed one to the other while maintaining
    the energy gap or localization; we will elaborate on that
    later. 
\end{quote}

\section{Anchor}
\label{sec:Anchor}

\begin{thebibliography}{1}
    \bibitem{kitaev-Periodic} Alexei Kitaev. Periodic table for
    topological insulators and superconductors.

\end{thebibliography}
\printnomenclature
\section{License}
The entire content of this work (including the source code
for TeX files and the generated PDF documents) by 
Hongxiang Chen (nicknamed we.taper, or just Taper) is
licensed under a 
\href{http://creativecommons.org/licenses/by-nc-sa/4.0/}{Creative 
Commons Attribution-NonCommercial-ShareAlike 4.0 International 
License}. Permissions beyond the scope of this 
license may be available at \url{mailto:we.taper[at]gmail[dot]com}.
\end{document}
